# Temporal Trend Analysis

## Setup

```{r setup, echo=FALSE}
source(here::here("R/_setup.R"))
load_or_save_latest_file(nfi_dataset_for_analysis, "load")
```

# TODO

-   Loop that every metric is plotted? Would require another nested loop within the first group function

-   Fix lower y label stuff

-   Region-based Analysis
    The idea of this section is to include a bootstrapping approach to calculate
    the uncertainty of the data. This comes at the advantage that we are taking together all data
    that we have on the trees and not subset for plots where there have been only a few trees
    of a given species.

# Functions
```{r}
tt_save_and_move_to_shiny <- function(
    my_tt, my_height = 7, my_width = 10
    ){
  
  # ______________________________________________________________________________
  # Make directories
  dir_figs <- get_todays_file_directory("figures")
  dir_figs_shiny <- get_todays_file_directory_shiny("figures")
  dir_rds_shiny <- get_todays_file_directory_shiny("rds")
  
  # ______________________________________________________________________________
  # Set Variables
  all_metrics <- get_ggplot_labs(return_available_vars = TRUE)
  all_files <- list()
  my_tt_name <- deparse(substitute(my_tt))
  
  # ______________________________________________________________________________
  # Start Loop
  if (!("plot_both_groups" %in% names(my_tt))) {
    # Loop for one group
    for (my_metric in all_metrics) {
      for (my_plot in c("plot_facet", "plot_allinone")) {
        
        my_p <- my_tt[[my_metric]][[my_plot]]
          
        my_file <- paste0(
            dir_figs, "/",
            my_tt_name, "-",
            my_metric, "-",
            my_plot,
            ".jpg"
          )
        
         my_file_shiny <- paste0(
            dir_figs_shiny, "/",
            my_tt_name, "-",
            my_metric, "-",
            my_plot,
            ".jpg"
          )
        
        ggsave(my_file,       my_p, height = my_height, width = my_width)
        ggsave(my_file_shiny, my_p, height = my_height, width = my_width)
        
        all_files <- append(all_files, my_file_shiny)
      }
    }
  } else {
  # Loop for two groups
    for (my_metric in all_metrics) {
        
      my_p <- my_tt[["plot_both_groups"]][[my_metric]]
        
      my_file <- paste0(
          dir_figs, "/",
          my_tt_name, "-",
          my_metric,
          ".jpg"
        )
      
       my_file_shiny <- paste0(
          dir_figs_shiny, "/",
          my_tt_name, "-",
          my_metric,
          ".jpg"
        )
      
      ggsave(my_file,       my_p, height = my_height, width = my_width)
      ggsave(my_file_shiny, my_p, height = my_height, width = my_width)
      
      all_files <- append(all_files, my_file_shiny)
    }
  }
  
  # ______________________________________________________________________________
  # Save path to plots
  path_to_plots <- all_files |> unlist()
  saveRDS(
    path_to_plots,
    paste0(dir_rds_shiny, "/path_to_plots_", my_tt_name,".rds")
    )
}
```

# Inputs
```{r}
my_min_trees_per_site <- 3
my_min_sites_per_year <- 5
```

# One Group
## Species
```{r}
tt_species <- 
  get_temporal_trends_for_1_group(
    df_in              = nfi_dataset_for_analysis,
    name_group_1       = "genus_lat",
    n_groups_1         = 12,
    min_trees_per_site = my_min_trees_per_site,
    min_sites_per_year = my_min_sites_per_year
  )

tt_save_and_move_to_shiny(tt_species, my_height = 7, my_width = 12)
```

## Height
```{r}
tt_height <- 
  get_temporal_trends_for_1_group(
    df_in              = nfi_dataset_for_analysis,
    name_group_1       = "height_class",
    n_groups_1         = 12,
    min_trees_per_site = my_min_trees_per_site,
    min_sites_per_year = my_min_sites_per_year
  )

tt_save_and_move_to_shiny(tt_height, my_height = 7, my_width = 12)
```

## GRECO
```{r}
tt_gre <- 
  get_temporal_trends_for_1_group(
    df_in              = nfi_dataset_for_analysis,
    name_group_1       = "gre",
    n_groups_1         = 12,
    min_trees_per_site = my_min_trees_per_site,
    min_sites_per_year = my_min_sites_per_year
)

tt_save_and_move_to_shiny(tt_gre, my_height = 7, my_width = 12)
```

# Two Groups

Note that patterns here are less clear and more variable because we require the data to have information on age and species.

## Species / Height

```{r}
tt_species_height <- 
  get_temporal_trends_for_2_groups(
    df_in              = nfi_dataset_for_analysis,
    name_group_1       = "genus_lat",
    name_group_2       = "height_class",
    n_groups_1         = 12,
    n_groups_2         = 10,
    min_trees_per_site = my_min_trees_per_site,
    min_sites_per_year = my_min_sites_per_year
  )

tt_height_species <- 
  get_temporal_trends_for_2_groups(
    df_in              = nfi_dataset_for_analysis,
    name_group_1       = "height_class",
    name_group_2       = "genus_lat",
    n_groups_1         = 12,
    n_groups_2         = 9,
    min_trees_per_site = my_min_trees_per_site,
    min_sites_per_year = my_min_sites_per_year
  )

tt_save_and_move_to_shiny(tt_species_height, my_height = 7, my_width = 12)
tt_save_and_move_to_shiny(tt_height_species, my_height = 7, my_width = 12)
```

## Species / GRECO

```{r}
tt_species_gre <- 
  get_temporal_trends_for_2_groups(
    df_in              = nfi_dataset_for_analysis,
    name_group_1       = "genus_lat",
    name_group_2       = "gre",
    n_groups_1         = 12,
    n_groups_2         = 9,
    min_trees_per_site = my_min_trees_per_site,
    min_sites_per_year = my_min_sites_per_year
  )

tt_gre_species <- 
  get_temporal_trends_for_2_groups(
    df_in              = nfi_dataset_for_analysis,
    name_group_1       = "gre",
    name_group_2       = "genus_lat",
    n_groups_1         = 12,
    n_groups_2         = 9,
    min_trees_per_site = my_min_trees_per_site,
    min_sites_per_year = my_min_sites_per_year
  )

tt_save_and_move_to_shiny(tt_species_gre, my_height = 7, my_width = 12)
tt_save_and_move_to_shiny(tt_gre_species, my_height = 7, my_width = 12)
```

## GRECO / Height

```{r}
tt_height_gre <- 
  get_temporal_trends_for_2_groups(
    df_in              = nfi_dataset_for_analysis,
    name_group_1       = "height_class",
    name_group_2       = "gre",
    n_groups_1         = 12,
    n_groups_2         = 9,
    min_trees_per_site = my_min_trees_per_site,
    min_sites_per_year = my_min_sites_per_year
  )

tt_gre_height <- 
  get_temporal_trends_for_2_groups(
    df_in              = nfi_dataset_for_analysis,
    name_group_1       = "gre",
    name_group_2       = "height_class",
    n_groups_1         = 12,
    n_groups_2         = 9,
    min_trees_per_site = my_min_trees_per_site,
    min_sites_per_year = my_min_sites_per_year
  )

tt_save_and_move_to_shiny(tt_height_gre, my_height = 7, my_width = 12)
tt_save_and_move_to_shiny(tt_gre_height, my_height = 7, my_width = 12)
```
